HERMES : Hybrid ERror-corrector Model with inclusion of External Signals for nonstationary fashion time series

### 2021-12-22 hermes_paper/hermes_ieee.tex
Corrections - Section 2

\newcommand{\ts}{y}
\newcommand{\fullts}{{\bf \ts}}
\newcommand{\tspred}{\hat{\ts}}
\newcommand{\stat}{f}
\newcommand{\statparam}{\theta_{predictor}}
\newcommand{\fullstat}{{\bf \stat}}
\newcommand{\lag}{h}
\newcommand{\window}{w}
\newcommand{\tswindow}{{\bf \ts}}
\newcommand{\meants}{\Bar{\ts}}
\newcommand{\rnnwindow}{{\bf \rnninput}}
\newcommand{\rnninput}{z}
\newcommand{\rnn}{\textsc{rnn}}
\newcommand{\rnnparam}{\theta_{corrector}}
\newcommand{\err}{err}
\newcommand{\errwindow}{{\bf \err}}
\newcommand{\rnnmodel}{\textsc{rnn}}
\newcommand{\ws}{w}
\newcommand{\fullws}{{\bf \ws}}
\newcommand{\wswindow}{{\bf \ws}}
\newcommand{\concatinput}{x}
\newcommand{\fullconcatinput}{{ \bf \concatinput}}
\newcommand{\numberts}{10000}
\newcommand{\threshold}{\eta}
\newcommand{\predictor}{\mathrm{RNN}_p}
\newcommand{\classifier}{\mathrm{RNN}_c}
\newcommand{\remainder}{r}
\newcommand{\hiddenregime}{U}

@@

\section{Hybrid model with external signals}
\label{sec:hybrid}
We introduce a new hybrid approach for time series forecasting  composed of two parts: a collection of per-time-series parametric models, the predictors, and a global error-corrector neural network train on all time series, the corrector. Per-time-series parametric models are used to learn local behaviour, to normalize sequences by removing trends and seasonality,  and to compute a first forecast. Then, gathering information of the first predictions and external variables, a recurrent neural networks is trained to correct the predictions provided by the first collection of per-time-series models.

@@

\subsection{Per-time-series predictors}
The time-series-specific predictors compute, for each sequence, a first $\lag$-ahead prediction based on the past. For all $1\leqslant n \leqslant N$, we note $\stat^n(.;\statparam^n)$ the $n$-th parametric model of the $n$-th sequence where $\statparam^n$ are  unknown parameters. Given the sequences $\{\ts^n_{1:T}\}_{1\leqslant n \leqslant N}$ and the estimated  parameters $\{\statparam^n\}_{1\leqslant n \leqslant N}$, the time-series-specific forecasts $\{\tspred^{pred,n}_{T+1:T+\lag|T}\}_{1\leqslant n \leqslant N}$ are, for all $n \in \{1,\ldots,N\}$, for all $i \in \{1,\ldots,\lag\}$,
$$
\tspred^{pred,n}_{T+i|T} = \stat^n(\ts^n_{1:T};\statparam^n)_i\,.
$$

During the M4 competition, the hybrid model of \cite{smyl2020hybrid} was based on a multiplicative exponential smoothing model as the time-series-specific predictor. However, on sporadic time series, this choice leads to poor result and instability. In this paper, a more general framework able to deal with any kind of per-time-series models is provided. In Section~\ref{sec:exp}, two versions of our framework are proposed. The first one use an  exponential smoothing as reference similar to the baseline \cite{smyl2020hybrid} and the second one use a TBATS model \cite{doi:10.1198/jasa.2011.tm09771} which provides better results as this parametric model includes  Fourier representations with time varying coefficients, and ARMA error correction.

In the specific application of fashion trends prediction, two main forms of standard weaknesses was detected and defined for an univariate parametric forecast as defined above. The first one is a weakness of definition, when the model is not well defined for its forecasting task. For instance, an additive model provide good predictions on time series with additive seasonality but show difficulties with multiplicative one. The second form, harder to detect and correct, is the weakness of information. For non stationary time series, huge changes of behaviours are not always predictable using the past of the sequence. In somes cases, these changes depend on external variables not considered by univariate parametric models. The difficulty is that the exact influence of external variables on the main signal is not mainly know. It is on this point that  the new HERMES framework shows its true potential: introduce a global RNN trained on all the time series and able to consider and leverage external signals. First forecast examples of this new hybrid model are given in Figure~\ref{fig:introexamples}.

@@

The second part of the model is a global RNN, trained on all the $N$ sequences to correct the weaknesses of the first per-time-series parametric models. This task requires a thorough data pre-processing for two main reasons. First, without a well-design input, the RNN will not be capable of understanding the initial collection of forecasts and predict their future errors. Second, as no assumption about the scale of our time series was made, inputs require a careful normalization before feeding them to the RNN.
Consider the following pre-processing.

Let $\window \in \mathbb{N}$ be the window size, usually this window is proportional to the forecast horizon $\window \propto \lag$. the RNN input is defined as the following  normalized, deseasonalized and rescaled sequence $\rnnwindow^n_T = \{\rnninput^{n}_{T-\window+i|T}\}_{1\leqslant i \leqslant w}$, where, for all $1\leqslant n \leqslant N$ and $1\leqslant i \leqslant w$, 
$$
\rnninput^{n,T}_{T-w+i|T} = \frac{\ts^n_{T-w+i} -\tspred^{pred,n}_{T+i|T}}{\meants^n_T}\,,\quad \meants^n_T = \frac{1}{w}\sum_{i = 1}^{w}\ts^n_{T-w+i}\,.
$$

Let $\rnn(.;\rnnparam)$ be the recurrent neural network model where $\rnnparam$ are  unknown parameters. Given the RNN input sequences $\{\rnnwindow^n_T\}_{1\leqslant n \leqslant N}$ and the global RNN estimated parameters $\rnnparam$, the error-corrector predictions $\{\tspred^{corr,n}_{T+1:T+\lag|T}\}_{1\leqslant n \leqslant N}$ are, for all $n \in \{1,\ldots,N\}$, for all $i \in \{1,\ldots,\lag\}$,
$$
\tspred^{corr,n}_{T+i|T} = \rnn(\rnnwindow^n_T;\rnnparam)_i \cdot \meants^n_T\ \,.
$$

Our hybrid model forecast is finally given, for all $1\leqslant n \leqslant N$ and all $i \in \{1,\ldots,\lag\}$, by:

\begin{align}
\label{eq:nows:full:model}
\tspred^n_{T+i|T}  &= \tspred^{pred,n}_{T+i|T} +  \tspred^{corr,n}_{T+i|T} \\
&= \stat^n(\ts^n_{1:T};\statparam^n)_i +  \rnn(\rnnwindow^n_T;\rnnparam)_i \cdot \meants^n_T\,.\nonumber
\end{align}

@@

Using well-fitted time-series-specific parametric models, the new hybrid network corrects the first form of weakness and achieves striking accuracy on the provided fashion dataset, see Table~\ref{tab:metricresults}. Then, to correct the second form of weakness, in addition to the $N$ target time series, $K \times N$ external sequences indexed from $0$ to $T$ are considered. For all $1\leqslant n \leqslant N$, $1\leqslant k \leqslant K$ and  $1\leqslant t \leqslant T$, let $\ws^{n,k}_t$ be the value of the $k$-th external sequence at time $t$ linked to the sequence $\fullts^n$. Let  $\fullws^n = \{\{\ws_t^{n,k}\}_{1\leqslant t \leqslant T}\}_{1\leqslant k \leqslant K}$ be all the values of the weak signals linked to the $n$-th sequence. In addition, let $\fullws^n_T = \{\{\ws_{T-w+i}^{n,k}\}_{1\leqslant i \leqslant \window}\}_{1\leqslant k \leqslant K}$ be only the last $\window$ terms of the sequence. Concatenating $ \rnnwindow^n_T$ and $\fullws^n_T$, a new input for the RNN is defined:   
\begin{align*}
\fullconcatinput^n_T &= \{\concatinput^n_{T-w+i|T}\}_{1\leqslant i \leqslant w}\\
&= \{\rnninput^n_{T-w+i|T}, \ws^{n,1}_{T-w+i},...,\ws^{n,K}_{T-w+i}\}_{1\leqslant i \leqslant w}\,.
\end{align*}
Finally, for all $1\leqslant n \leqslant N$ and for all $i \in \{1,\ldots,\lag\}$ the final prediction becomes:
%$$
%\tspred^n_{T+i|T} 
%= \etspred^{ets,n}_{T+i|T} +  \rnnmodel(\fullconcatinput^n_T)_{T+i} \times \meants^n_T
%$$
\begin{align}
\label{eq:withws:full:model}
\tspred^n_{T+i|T}  &= \tspred^{pred,n}_{T+i|T} +  \tspred^{corr,n}_{T+i|T}\\
& = \stat^n(\ts^n_{1:T};\statparam^n)_i +  \rnn(\fullconcatinput^n_T;\rnnparam)_i \cdot \meants^n_T\,.\nonumber
\end{align}
An illustration of the proposed architecture is displayed in Figure~\ref{fig:architecture}.

>>>>

\newcommand{\ts}{y}
\newcommand{\fullts}{{\bf \ts}}
\newcommand{\tspred}{\widehat{\ts}}
\newcommand{\stat}{f}
\newcommand{\statparam}{\theta_{predictor}}
\newcommand{\fullstat}{{\bf \stat}}
\newcommand{\lag}{h}
\newcommand{\window}{w}
\newcommand{\tswindow}{{\bf \ts}}
\newcommand{\meants}{\Bar{\ts}}
\newcommand{\rnnwindow}{{\bf \rnninput}}
\newcommand{\rnninput}{z}
\newcommand{\rnn}{\textsc{rnn}}
\newcommand{\rnnparam}{\theta_{corrector}}
\newcommand{\err}{err}
\newcommand{\errwindow}{{\bf \err}}
\newcommand{\rnnmodel}{\textsc{rnn}}
\newcommand{\ws}{w}
\newcommand{\fullws}{{\bf \ws}}
\newcommand{\wswindow}{{\bf \ws}}
\newcommand{\concatinput}{x}
\newcommand{\fullconcatinput}{{ \bf \concatinput}}
\newcommand{\numberts}{10000}
\newcommand{\threshold}{\eta}
\newcommand{\predictor}{\mathrm{RNN}_p}
\newcommand{\classifier}{\mathrm{RNN}_c}
\newcommand{\remainder}{r}
\newcommand{\hiddenregime}{U}

@@

\section{Hybrid model with external signals}
\label{sec:hybrid}
We introduce a new hybrid approach for time series forecasting  composed of two parts: a collection of per-time-series parametric models, and a global error-corrector neural network train on all time series. Per-time-series parametric models are used to learn local behaviours, to normalize sequences by removing trends and seasonality,  and to compute a first forecast. Then, gathering information of the first predictions and external variables, a recurrent neural networks is trained to correct the predictions provided by the first collection of per-time-series models.

@@

\subsection{Per-time-series predictors}
The time-series-specific predictors compute, for each sequence, a first $\lag$-ahead prediction based on the past. For all $1\leqslant n \leqslant N$, we note $\stat^n(.;\statparam^n)$ the $n$-th parametric model of the $n$-th sequence where $\statparam^n$ are  unknown parameters. Given the sequences $\{\ts^n_{1:T}\}_{1\leqslant n \leqslant N}$ and the estimated  parameters $\{\statparam^n\}_{1\leqslant n \leqslant N}$, the time-series-specific forecasts $\{\tspred^{pred,n}_{T+1:T+\lag|T}\}_{1\leqslant n \leqslant N}$ are, for all $n \in \{1,\ldots,N\}$, for all $i \in \{1,\ldots,\lag\}$,
$$
\tspred^{pred,n}_{T+i|T} = \stat^n(\ts^n_{1:T};\statparam^n)_i\,.
$$
During the M4 competition, the hybrid model of \cite{smyl2020hybrid} was based on a multiplicative exponential smoothing model as the time-series-specific predictor. However, on sporadic time series, this choice leads to poor results and instability. In this paper, a more general framework able to deal with any kind of per-time-series models is provided. In Section~\ref{sec:exp}, two versions of our framework are proposed. The first one is based on an  exponential smoothing as a reference similar to the baseline \cite{smyl2020hybrid} and the second one use a TBATS model \cite{doi:10.1198/jasa.2011.tm09771} which provides better results as this parametric model includes  Fourier representations with time varying coefficients, and ARMA error correction.

In the specific application of fashion trends prediction, two main forms of standard weaknesses were detected and defined for univariate parametric forecasts. The first one is a weakness of definition, when the model is not well defined for its forecasting task. For instance, an additive model provides good predictions on time series with additive seasonality but fails at handling multiplicative seasonalities. The second form, harder to detect and correct, is the weakness of information. For non stationary time series, huge changes of behaviours are not always predictable using the past of the sequence. In somes cases, these changes depend on external variables not considered by univariate parametric models. The difficulty is that the exact influence of external variables on the main signal is mostly unknown. It is on this point that  the new HERMES framework shows its true potential: introduce a global RNN trained on all the time series and able to consider and leverage external signals. First forecast examples of this new hybrid model are given in Figure~\ref{fig:introexamples}.

@@

The second part of the model is a global RNN, trained on all the $N$ sequences to correct the weaknesses of the first per-time-series parametric models. This task requires a thorough data pre-processing as recurrent neural networks training is highly sensitive to the scale of the data and requires well-designed inputs. Since no assumption about the scale of our time series was made, inputs require a careful normalization before being fed to the RNN.

Let $\window \in \mathbb{N}$ be the window size, usually this window is proportional to the forecast horizon $\window \propto \lag$. the RNN input is defined as the following  normalized, deseasonalized and rescaled sequence $\rnnwindow^n_T = \{\rnninput^{n}_{T-\window+i|T}\}_{1\leqslant i \leqslant w}$, where, for all $1\leqslant n \leqslant N$ and $1\leqslant i \leqslant w$, 
$$
\rnninput^{n,T}_{T-w+i|T} = \frac{\ts^n_{T-w+i} -\tspred^{pred,n}_{T+i|T}}{\meants^n_T}\,,\quad \meants^n_T = \frac{1}{w}\sum_{i = 1}^{w}\ts^n_{T-w+i}\,.
$$
Let $\rnn(.;\rnnparam)$ be the recurrent neural network model where $\rnnparam$ are  unknown parameters. Given the RNN input sequences $\{\rnnwindow^n_T\}_{1\leqslant n \leqslant N}$ and the global RNN estimated parameters $\rnnparam$, the error-corrector predictions $\{\tspred^{corr,n}_{T+1:T+\lag|T}\}_{1\leqslant n \leqslant N}$ are, for all $n \in \{1,\ldots,N\}$, for all $i \in \{1,\ldots,\lag\}$,
$$
\tspred^{corr,n}_{T+i|T} = \rnn(\rnnwindow^n_T;\rnnparam)_i \cdot \meants^n_T\ \,.
$$
Our hybrid model forecast is, for all $1\leqslant n \leqslant N$ and all $i \in \{1,\ldots,\lag\}$,
\begin{align}
\label{eq:nows:full:model}
\tspred^n_{T+i|T}  &= \tspred^{pred,n}_{T+i|T} +  \tspred^{corr,n}_{T+i|T} \\
&= \stat^n(\ts^n_{1:T};\statparam^n)_i +  \rnn(\rnnwindow^n_T;\rnnparam)_i \cdot \meants^n_T\,.\nonumber
\end{align}

@@

Using well-fitted time-series-specific parametric models, the new hybrid network corrects the first form of weakness and achieves striking accuracy on the provided fashion dataset, see Table~\ref{tab:metricresults}. Then, to correct the second form of weakness, in addition to the $N$ target time series, $K \times N$ external sequences indexed from $0$ to $T$ are considered. For all $1\leqslant n \leqslant N$, $1\leqslant k \leqslant K$ and  $1\leqslant t \leqslant T$, let $\ws^{n,k}_t$ be the value of the $k$-th external sequence at time $t$ associated with the sequence $\fullts^n$. Let  $\fullws^n = \{\{\ws_t^{n,k}\}_{1\leqslant t \leqslant T}\}_{1\leqslant k \leqslant K}$ be all the values of the weak signals associated with the $n$-th sequence. In addition, let $\fullws^n_T = \{\{\ws_{T-w+i}^{n,k}\}_{1\leqslant i \leqslant \window}\}_{1\leqslant k \leqslant K}$ be only the last $\window$ terms of the sequence. Concatenating $ \rnnwindow^n_T$ and $\fullws^n_T$, a new input for the RNN is defined:   
\begin{align*}
\fullconcatinput^n_T &= \{\concatinput^n_{T-w+i|T}\}_{1\leqslant i \leqslant w}\\
&= \{\rnninput^n_{T-w+i|T}, \ws^{n,1}_{T-w+i},...,\ws^{n,K}_{T-w+i}\}_{1\leqslant i \leqslant w}\,.
\end{align*}
Finally, for all $1\leqslant n \leqslant N$ and for all $i \in \{1,\ldots,\lag\}$ the final prediction becomes:
%$$
%\tspred^n_{T+i|T} 
%= \etspred^{ets,n}_{T+i|T} +  \rnnmodel(\fullconcatinput^n_T)_{T+i} \times \meants^n_T
%$$
\begin{align}
\label{eq:withws:full:model}
\tspred^n_{T+i|T}  &= \tspred^{pred,n}_{T+i|T} +  \tspred^{corr,n}_{T+i|T}\\
& = \stat^n(\ts^n_{1:T};\statparam^n)_i +  \rnn(\fullconcatinput^n_T;\rnnparam)_i \cdot \meants^n_T\,.\nonumber
\end{align}
An illustration of the proposed  model is displayed in Figure~\ref{fig:architecture}.
